---
title: "taylor-swift-music-analysis"
author: "Malina Freeman"
date: "2024-02-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## R Markdown - Taylor Swift Music Analysis

#### Question: Do All Taylor Swift Songs Actually Sound the Same?

Imports the data set from my files and views it

```{r}
taylor_album_songs <- read.csv("C:/Users/malin/Downloads/taylor_album_songs.csv")
View(taylor_album_songs)
```

Use the glimpse function to get a better idea of categories of data set

```{r}
glimpse(taylor_album_songs)
```

Finding average of various song metrics

```{r}
mean(taylor_album_songs$danceability, na.rm = TRUE)
mean(taylor_album_songs$acousticness, na.rm = TRUE)
mean(taylor_album_songs$loudness, na.rm = TRUE)
mean(taylor_album_songs$energy, na.rm = TRUE)
```
Summarizing each category for data

```{r}
summary(taylor_album_songs)
```

Removing null values in categories of interest

```{r}
taylor_album_songs$danceability <-ifelse(is.na(taylor_album_songs$danceability),
    median(taylor_album_songs$danceability, na.rm = TRUE),
    taylor_album_songs$danceability)
taylor_album_songs$loudness <-ifelse(is.na(taylor_album_songs$loudness),
    median(taylor_album_songs$loudness, na.rm = TRUE),
    taylor_album_songs$loudness)
taylor_album_songs$acousticness <-ifelse(is.na(taylor_album_songs$acousticness),
    median(taylor_album_songs$acousticness, na.rm = TRUE),
    taylor_album_songs$acousticness)
taylor_album_songs$energy <-ifelse(is.na(taylor_album_songs$energy),
    median(taylor_album_songs$energy, na.rm = TRUE),
    taylor_album_songs$energy)
taylor_album_songs$instrumentalness <-ifelse(is.na(taylor_album_songs$instrumentalness),
    median(taylor_album_songs$instrumentalness, na.rm = TRUE),
    taylor_album_songs$instrumentalness)
taylor_album_songs$duration_ms <-ifelse(is.na(taylor_album_songs$duration_ms),
    median(taylor_album_songs$duration_ms, na.rm = TRUE),
    taylor_album_songs$duration_ms)
taylor_album_songs$explicit <-ifelse(is.na(taylor_album_songs$explicit),
    median(taylor_album_songs$explicit, na.rm = TRUE),
    taylor_album_songs$explicit)
taylor_album_songs$valence <-ifelse(is.na(taylor_album_songs$valence),
    median(taylor_album_songs$valence, na.rm = TRUE),
    taylor_album_songs$valence)
taylor_album_songs$tempo <-ifelse(is.na(taylor_album_songs$tempo),
    median(taylor_album_songs$tempo, na.rm = TRUE),
    taylor_album_songs$tempo)
summary(taylor_album_songs)
```

Selecting the specific columns of interest and dropping other columns

```{r}
taylor_album_songs <- select(taylor_album_songs, valence, explicit, danceability, acousticness, instrumentalness, energy,loudness, album_name, tempo)
summary(taylor_album_songs)
```

Grouping by album and printing the mean of different categories for songs in that album

```{r}
taylor_album_songs %>%
  group_by(album_name) %>%
  summarise(mean(energy),
            mean(valence),
            mean(acousticness),
            mean(danceability))

```

Showing a histogram of the energy of different albums

```{r}
ggplot(taylor_album_songs, aes(x = energy,
                               color = album_name)) +
  geom_histogram() 

```
Trying a scatter plot correlating valence (x-axis) and energy (y-axis)

```{r}
ggplot(taylor_album_songs, aes(x = valence,
                               y = energy,
                               color = album_name)) +
  geom_point()
```
#### Use Statistical Analysis to Determine if All Songs Sound the Same 

```{r}
summary(taylor_album_songs)
```
First Creating a New Data frame based on song dance-ability
```{r}
weight <- c(taylor_album_songs$danceability)
group <- c(taylor_album_songs$album_name)
taylor_danceability <- data.frame(weight, group)
print(taylor_danceability)
```


```{r}

p <- ggboxplot(taylor_danceability, x = "group", y = "weight")
p + rotate_x_text()
```

Statistical Test measuring difference between medians of the groups (e.g. albums)
```{r}
kruskal.test(weight ~ group, data = taylor_danceability)
```
Low p-value indicates we can reject the null hypothesis that the medians of all the groups are the same. There is a statistically significant difference between at least two groups for dance-ability!

#### Apply this Method for Other Categories
Note: valence = how positive a song sounds according to Spotify
```{r}
weight <- c(taylor_album_songs$valence)
group <- c(taylor_album_songs$album_name)
taylor_valence <- data.frame(weight, group)
print(taylor_valence)
```
```{r}

p <- ggboxplot(taylor_valence, x = "group", y = "weight")
p + rotate_x_text()
```

There is not as much of a statistically significant difference, but p-value is still less than 0.05 
```{r}
kruskal.test(weight ~ group, data = taylor_valence)
```
```{r}
weight <- c(taylor_album_songs$tempo)
group <- c(taylor_album_songs$album_name)
taylor_tempo <- data.frame(weight, group)
print(taylor_tempo)
```

```{r}

p <- ggboxplot(taylor_tempo, x = "group", y = "weight")
p + rotate_x_text()
```

NOT a statistically significant difference!
```{r}
kruskal.test(weight ~ group, data = taylor_tempo)
```
```{r}
weight <- c(taylor_album_songs$energy)
group <- c(taylor_album_songs$album_name)
taylor_energy <- data.frame(weight, group)
print(taylor_energy)
```

```{r}

p <- ggboxplot(taylor_energy, x = "group", y = "weight")
p + rotate_x_text()
```

```{r}
kruskal.test(weight ~ group, data = taylor_energy)
```